---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(scipen = 1, digits = 2)
```

# gamdiffs

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/gamdiffs)](https://CRAN.R-project.org/package=gamdiffs)
[![R-CMD-check](https://github.com/sempwn/gamdiffs/workflows/R-CMD-check/badge.svg)](https://github.com/sempwn/gamdiffs/actions)
<!-- badges: end -->

The goal of gamdiffs is to provide a set of convenience functions for the estimation
and associated standard errors / confidence intervals for a range of Generalized
Additive Models. The particular motivation is to provide tools for estimating the 
difference in outcome under certain counterfactual scenarios for models relevant
to population public health.

Current implemented confidence interval estimates are

* delta method
* posterior sampling
* Efron bootstrap sampling


## Installation

You can install the released version of gamdiffs from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("gamdiffs")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("sempwn/gamdiffs")
```
## Example

This basic example generates some data and fit a thin-plate spline to a time
variable (`x`) with outcome `y` which is Poisson-distributed. The sum of the outcome
between time 0 to 20 is then compared to the outcome between time 21 to time 40 
and the resulting expected difference with associated confidence intervals is 
provided

```{r example}
library(gamdiffs)
library(mgcv)
## basic usage
res <- gamdiffs:::create_random_data()
m <- mgcv::gam(y ~ s(x), data = res, family = poisson)
baseline_data <- dplyr::tibble(x = 0:20)
counter_data <- dplyr::tibble(x = 21:40)
test_diffs <- calc_sum_counterfactual_gam(m, baseline_data,
  counter_data = counter_data,
  ci = 0.95
)
```

The resulting expected sum difference between the `baseline_data` and `counter_data`
are

```{r print result}
print(test_diffs)
```

The default setting is to estimate the confidence interval from the delta 
method. The resulting confidence interval from sampling of the posterior (with
improper priors) can be calculated as

```{r example 2}

post_diffs <- calc_sum_counterfactual_gam(m, 
  baseline_data,
  use_post = TRUE,
  nrep = 1000,
  counter_data = counter_data,
  ci = 0.95
)
print(post_diffs)
```


## Code of Conduct
  
  Please note that the gamdiffs project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
